Parallel multi-objective optimization using co-evolutionary islands and search space overlapping

### 2016-12-13 hpmoon-coor.tex
Explaining the overlapped stuff. Closes #1

\begin{abstract}
Multi-objective optimization methods have gained interest in the last years because the performance they can achieve solving real-world problems. Many of these problems usually require a high-dimensional decision space, and multi-objective Evolutionary Algorithms have been successfully used in the past. Due to the large computational cost these algorithms require, different parallelization and distribution methods have been proposed, with different parallel and distributed computer architectures to cope with them. One of the most extended methods are the distributed co-evolutionary algorithms, where each processor deals only with a part of the decision space, that is, the individuals belonging to the same sub-population (island) explore the same subset of decision variables.

In this paper, a new method to automatically adapt the overlapping sections of the individuals has been compared to other collaborative evolutionary methods. Different number of islands, chromosome sizes and problems have been used. The analysis of the obtained experimental results, by using different metrics, shows that coevolutionary approaches can provide statistically significant improvements with respect to the base algorithm, being the method to automatically adapt the overlapping size the one to obtain the best performance during the same execution time. Results also show that the relation of the number of islands (subpopulations) to the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distribute the decision variables.
\end{abstract}

@@

In our preliminary work in \citep{Garcia16hpmoon} we tried to demonstrate that using overlapped sections of the chromosome in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time when the number of islands increases. We discovered that the most adequate overlapping size may vary depending on a relation between the number of islands and the individual size. Moreover, the experiments were conducted in a mono-processor and synchronous island model. This motivates us to continue this research line, but this time, proposing a new method to automatically adapt the size of the overlapped sections, but also, performing all the experiments in a real cluster with up to 128 nodes, with extra number of islands.

>>>>

\begin{abstract}
Multi-objective optimization methods have gained interest in the last years because the performance they can achieve solving real-world problems. Many of these problems usually require a high-dimensional decision space, and multi-objective Evolutionary Algorithms have been successfully used in the past. Due to the large computational cost these algorithms require, different parallelization and distribution methods have been proposed, with different parallel and distributed computer architectures to cope with them. One of the most extended methods are the distributed co-evolutionary algorithms, where each processor deals only with a part of the decision space, that is, the individuals belonging to the same sub-population (island) explore the same subset of decision variables. These subsets are usually disjoint sets.

In our previous work we demonstrated that extending the length of these subsets, and therefore, sharing some search space between islands, implies an improvement in the results. In this paper, a new method to automatically adapt the size of the overlapping, or shared, sections of the individuals has been compared to other collaborative evolutionary techniques. Different number of islands, chromosome sizes and problems have been used. The analysis of the obtained experimental results, by using different metrics, shows that coevolutionary approaches can provide statistically significant improvements with respect to the base algorithm, being the method to automatically adapt the overlapping size the one to obtain the best performance during the same execution time. Results also show that the relation of the number of islands (subpopulations) to the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distribute the decision variables.
\end{abstract}

@@

In our preliminary work in \citep{Garcia16hpmoon} we tried to demonstrate that, instead focus every island in a disjoint subset of the chromosome, the usage of overlapped (shared) sections of the chromosome in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time when the number of islands increases. We discovered that the most adequate overlapping size may vary depending on a relation between the number of islands and the individual size. Moreover, the experiments were conducted in a mono-processor and synchronous island model. This motivates us to continue this research line, but this time, proposing a new method to automatically adapt the size of the overlapped sections, but also, performing all the experiments in a real cluster with up to 128 nodes, with extra number of islands.
